%!TEX root =  ../main.tex

\mychapters{Functions}{functions}{\chapdir/pics/Mountain-type_Woodland_Caribou}


Functions are the central objects of investigation in many subdivision of mathematics today, and
are very effective means of representing real-life phenomena. For example, in any given year, 
there is one population of
deer in a certain forest.  Would it make sense for a forest ranger to answer you two numbers when
you ask how many \textit{Rangifer tarandus} (caribou) exist in a given range, at a certain time?


\begin{figure}[h]
\centering
\begin{tikzpicture}
    \begin{axis}%
    [
        %grid=major,     
        xmin=0,
        xmax=12,
        axis x line=bottom,
        ytick={0,10,...,100},
        ymax=110
       % axis y line=middle,
    ]
        \addplot%
        [
            blue,%
            mark=none,
            samples=100,
            domain=0:12,
        ]
        (x,{1+100/(1+exp(-x+5))});
    \end{axis}
\end{tikzpicture}
\caption{Population of caribou over time}
\end{figure}

How do we model phenomena in the world?  Is it reasonable to expect
caribou population to follow a curve such as the above figure?  How can we 
extract information from a graph like this?  If actual measurements diverge some
from the expectation established by the curve, are they \emph{wrong}?  Are functions
better tools than relations?  What else could there be?

\newpage
\chapterminitoc

This chapter is entitled functions.  You have been studying functions for several years now,
and should be familiar with function notation.  It is also possible to present functions 
graphically, or in a table.  Lastly, we may write or speak a verbal description of a function,
but this requires technical jargon, with which you are just starting to become familiar.

Hopefully, your time in algebra classes until now has helped you achieve a basic familiarity
with certain elementary or parent functions.  You will need to recognize them in each
of the four ways.  These prototypical versions can be moved or stretched as needed.  They
can also be mirrored or reflected.  This should naturally lead to the observation that some
graphs are unchanged under certain types of reflection.

Lastly, we will introduce the rich and vast field of taking numerical data and generating
an algebraic equation, called regression.  


%								1 - 1
\newpage
\invisiblesection{Representations}
\marginlessinput{\chapdir/0101p}
\newpage
\input{\chapdir/0101}
\newpage
\subsection{Exercises}
\marginlesspdf{\chapdir/0101xA.pdf}
\marginlesspdf{\chapdir/0101xB.pdf}
\marginlesspdf{\chapdir/0101xC.pdf}



%								1 -2
\newpage
\invisiblesection{Types}
\subsection{Pick x, Find y}
\marginlesspdf{ch01/0102p.pdf}
\newpage
\input{ch01/0102}
\newpage
\subsection{Exercises}
in Kuta
~\vfill


%								1 - 3
\newpage
\invisiblesection{Translations and Dilations}
\subsection{Multiplying and Adding Functions}
\marginlesspdf{ch01/0103pA.pdf}
%\newpage
\marginlesspdf{ch01/0103pB.pdf}
\newpage
\input{ch01/0103}
\newpage
\subsection{Exercises}
in Kuta
~\vfill


%								1 - 4
\newpage
\invisiblesection{Reflections and Symmetries}%
\marginlessinput{ch01/0104p}
\newpage
\input{ch01/0104}
\newpage
\subsection{Exercises}
in Kuta


%								1 - 5
\newpage
\invisiblesection{Regressions}
\subsection{Graphical Patterns}
\marginlesspdf{ch01/0105pA.pdf}
\newpage
\marginlesspdf{ch01/0105pB.pdf}
\input{ch01/0105}
\newpage
\input{\chapdir/0105x}




%								1 - 6
\newpage
\section{Review}
\input{ch01/0106}
