\renewcommand{\columnseprule}{1.5pt}
\begin{multicols*}{2}
\noindent
\rule[0.5\baselineskip]{0.5\textwidth}{1pt}

\noindent
\subsection{Families of Functions}
\noindent
1. Sketch the graph of $f(x)=\frac{1}{2}x-3$ and label the coordinates of the $x$ and $y$ intercepts.
Describe the obvious etymology of the adjective \textbf{linear} in a whole sentence.

\vspace{3cm}
\noindent
2. Sketch the \textbf{quadratic} function $f(x)=-x^2+2x+8$, labelling its intercepts, \textbf{vertex},
and axis of symmetry.  Look up the etymology of the word ``quadratic'' and explain why the 
Latin root for `four' is in our word for an equation with a power of two.

\vspace{4cm}

\noindent
3. Sketch the graph of the \textbf{power} function $f(x)=2x^\frac{3}{4}$.  When is the next lattice point
after (1,2)?

\vspace{3cm}
\noindent
4. Sketch the graph of the \textbf{exponential} function $f(x)=5\cdot{}.9^x$.  Why do you think it is called
an ``exponential'' function and how does it contrast with the previous problem (algebraically and graphically)?

\vspace{3cm}
\noindent
5. Sketch the graph of the \textbf{logarithmic} function $f(x)=2\ln{2x)-6$.  Does the $y$ value ever exceed 0?
If there a vertical asymptote at $x=0$?  Why does you calculator stop graphing around (0.15,-8.38)?

\vspace{3cm}
\noindent
6. Not all numbers make sense, either as outputs or inputs to your function.  Conjecture
a reason \textbf{domain} and \textbf{range} and tell how you arrived at the intervals you did.
Record them below.

\vspace{3cm}
\noindent
7. When using a \textbf{mathematical model}, there are two words to describe going beyond
recorded data:  \textbf{interpolate} and \textbf{extrapolate}.  Which prefix means `within' and which means
`without'?  Which word is appropriate to describe finding the missing data point you skipped over from
your observations?  Which would is appropriate to describe finding how long 100 objects 
would be?  Explain your choice.

\vspace{5cm}

8. Describe what you think the point of this problem set is, using technical vocabulary in complete
sentences.
\end{multicols*}